---
title: "Hass Avocado Analysis"
author: "Jeremy Meguira"
date: "October 9, 2018"
output: 
  html_document: 
    fig_height: 10
    fig_width: 10
    toc: yes
    toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## Hass Avocado Analysis
This dataset was pulled from:
https://www.kaggle.com/neuromusic/avocado-prices/home

It contains historical data on avocado prices and sales volume in multiple US
markets.

Some questions I want to look at in this analysis:

* How do the prices of avocados vary by region?
* What happened to avocado prices and consumption in 2017?
* How does consumption/pricing change based on organic/conventional avocados?

If any other interesting questions come up I will look into them as well.


```{r echo = FALSE}

library(ggplot2)
library(knitr)
library(GGally)
library(gridExtra)

```

## Load and Examine Dataset

```{r echo = FALSE, Loading_Data}

av <- read.csv('avocado.csv')

```

Variable descriptions are provided by the dataset poster:

* Date - The date of the observation
* AveragePrice - the average price of a single avocado
* type - conventional or organic
* year - the year
* Region - the city or region of the observation
* Total Volume - Total number of avocados sold
* 4046 - Total number of avocados with PLU 4046 sold / small hass avocados
* 4225 - Total number of avocados with PLU 4225 sold / large hass avocados
* 4770 - Total number of avocados with PLU 4770 sold / x-large hass avocados

## Clean

Check for duplicate rows

```{r echo = FALSE}

sum(duplicated(av))

```

Check for missing values

```{r echo = FALSE}

sum(is.na(av))

```

Looks like a pretty clean data set. No missing values or duplicated rows. Now 
let's look into the individual variables a bit

```{r echo = FALSE}

str(av)
 
```

The  year variable is redundant, as that information is already contained in the
date variable, so we can drop it.

```{r echo = FALSE}

#https://stackoverflow.com/a/4605268

drops <- c('year')
av <- av[ , !names(av) %in% drops]
str(av)


```

As we can see, date is currently of Factor type which is usually reserved for
categorical data. I will change it to a date object

```{r echo = FALSE}

av$Date <- as.POSIXct(av$Date)
str(av$Date)

```

Since we know the price of the average avocado at the date of the observation
and the total # of avocados sold, we can create a new variable gross_sales to
show the income for that week at each location purely from avocado sales.

Summary statistics for gross sales:

```{r echo = FALSE}

av$gross_sales <- av$Total.Volume * av$AveragePrice 
summary(av$gross_sales)

```

After some exploration, I realized that it didn't make sense to chart
organic and conventional avocados together and that I had to breakup the dataset
by region in order to create graphs that made sense. I will create a subset
for the TotalUS region, the larger regions that encompass more than one state 
(e.g.: 'Midsouth', 'Northeast'), and individual cities. These will be subset one
more time by avocado type.

```{r echo = FALSE}

unique(av$region)

```


```{r echo = FALSE}

av_total_org <- subset(av, av$region == "TotalUS" & av$type == "organic")
av_total_con <- subset(av, av$region == "TotalUS" & av$type == "conventional")

region_list <- list('GreatLakes','Midsouth', 'Northeast','NorthernNewEngland','West', 'SouthCentral', 'Southeast','Plains', 'West')
av_regional_org <- subset(av, av$region %in% region_list & av$type == "organic")
av_regional_con <- subset(av, av$region %in% region_list & av$type == "conventional")

region_list <- list('GreatLakes','Midsouth', 'Northeast','NorthernNewEngland','West', 'SouthCentral', 'Southeast','Plains', 'West', 'TotalUS')
av_city_org <- subset(av, !(av$region %in% region_list) & av$type == "organic")
av_city_con <- subset(av, !(av$region %in% region_list) & av$type == "conventional")

```



## Explore



```{r echo = FALSE}

print('Total organic avocado sales:')
sum(av_total_org$Total.Volume)
print('Total conventional avocado sales:')
sum(av_total_con$Total.Volume)

print('Gross sales for organic avocados:')
sum(av_total_org$gross_sales)
print('Gross sales for conventional avocados:')
sum(av_total_con$gross_sales)

print('Average price of organic avocado:')
mean(av_total_org$AveragePrice)
print('Average price of conventional avocado:')
mean(av_total_con$AveragePrice)

```

Avocado total price distribution:

```{r echo = FALSE}

ap1 <- ggplot(aes(y = AveragePrice), data = av) +
  geom_boxplot(fill = 'orange') + 
  ylab('Average Avocado Price')

ap2 <- ggplot(aes(x = AveragePrice), data = av) +
  geom_histogram(fill = 'orange')

grid.arrange(ap1, ap2, ncol = 1)

```

Price distribution by avocado type:

```{r echo = FALSE}

ap1_org <- ggplot(aes(y = av_total_org$AveragePrice), data = av_total_org) +
  geom_boxplot(fill = 'orange') + 
  ylab('Average Organic Avocado Price')

ap2_org <- ggplot(aes(x = av_total_org$AveragePrice), data = av_total_org) +
  geom_histogram(fill = 'orange')

ap1_con <- ggplot(aes(y = av_total_con$AveragePrice), data = av_total_con) +
  geom_boxplot(fill = 'orange') + 
  ylab('Average Conventional Avocado Price')

ap2_con <- ggplot(aes(x = av_total_con$AveragePrice), data = av_total_con) +
  geom_histogram(fill = 'orange')

grid.arrange(ap1_org, ap2_org, ap1_con, ap2_con, ncol = 2) 

```

TotalUS Plots

```{r}

```


Regional Plots

```{r}

```


City Plots Conventional

```{r echo = FALSE}

ggplot(aes(x = Date, y = Total.Volume), data = av_total_con) + 
  geom_point(stat = 'identity')

ggplot(aes(x = Date, y = Total.Volume), data = av_total_org) +
  geom_point(stat = 'identity')

ggplot(aes(x = Date, y = Total.Volume, color = 'orange' ), data = av_city_con) + 
  geom_point(stat = 'summary', fun.y = mean) +
  stat_smooth(method = 'lm', color = 'blue') +
  coord_cartesian(ylim = c(300000, quantile(av_city_con$Total.Volume,.9)))
  #scale_x_datetime(limits = c(as.POSIXct("2017-01-01"), as.POSIXct("2018-01-01")))

ggplot(aes(x = Date, y = AveragePrice, color = 'orange'), data = av_city_con) + 
  geom_point(stat = 'summary', fun.y = mean) +
  stat_smooth(method = 'lm', color = 'blue')
  #coord_cartesian(ylim = c(300000, quantile(av_city_con$Total.Volume,.9)))

ggplot(aes(x = Date, y = gross_sales, color = 'orange'), data = av_city_con) +
  geom_point(stat = 'summary', fun.y = mean) +
  stat_smooth(method = 'lm', color = 'blue')

ggplot(aes(x = Date, y = AveragePrice, color = ))
```

```{r echo = FALSE}


```

