---
title: "Hass Avocado Analysis"
author: "Jeremy Meguira"
date: "October 9, 2018"
output: 
  html_document: 
    fig_height: 10
    fig_width: 10
    toc: yes
    toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
options(scipen=999)
```

## Hass Avocado Analysis
This dataset was pulled from:
https://www.kaggle.com/neuromusic/avocado-prices/home

It contains historical data on avocado prices and sales volume in a variety of
US markets.

Some questions I want to look at in this analysis:

* What happened to avocado prices and consumption in 2017?
* How do the prices of avocados vary by region?



```{r echo = FALSE}

library(ggplot2)
library(knitr)
library(GGally)
library(gridExtra)

```

## Load and Examine Dataset

```{r echo = FALSE, Loading_Data}

av <- read.csv('avocado.csv')

```

Variable descriptions are provided by the dataset poster:

* Date - The date of the observation
* AveragePrice - the average price of a single avocado
* type - conventional or organic
* year - the year
* Region - the city or region of the observation
* Total Volume - Total number of avocados sold
* 4046 - Total number of avocados with PLU 4046 sold / small hass avocados
* 4225 - Total number of avocados with PLU 4225 sold / large hass avocados
* 4770 - Total number of avocados with PLU 4770 sold / x-large hass avocados

## Clean

Check for duplicate rows

```{r echo = FALSE}

sum(duplicated(av))

```

Check for missing values

```{r echo = FALSE}

sum(is.na(av))

```

Looks like a pretty clean data set. No missing values or duplicated rows. Now 
let's look into the individual variables a bit

```{r echo = FALSE}

str(av)
 
```

The  year variable is redundant, as that information is already contained in the
date variable, so we can drop it.

```{r echo = FALSE}

#https://stackoverflow.com/a/4605268

drops <- c('year')
av <- av[ , !names(av) %in% drops]
str(av)


```

As we can see, date is currently of Factor type which is usually reserved for
categorical data. I will change it to one of r's date objects, POSIXct

```{r echo = FALSE}

av$Date <- as.POSIXct(av$Date)
str(av$Date)

```

Since we know the price of the average avocado at the date of the observation
and the total # of avocados sold, we can create a new variable gross_sales to
show the income for that week at each location purely from avocado sales.

Summary statistics for gross sales:

```{r echo = FALSE}

av$gross_sales <- av$Total.Volume * av$AveragePrice 
summary(av$gross_sales)

```

After some exploration, I realized that it didn't make sense to chart organic
and conventional avocados together as their sales numbers are orders of
magnitude apart. I broke up the dataset by region in order to create graphs that
made sense. I will create a subset for the TotalUS region, the larger regions
that encompass more than one state (e.g.: 'Midsouth', 'Northeast'), and
individual cities. Below is a list of the unique regions included in this
dataset.

```{r echo = FALSE}

unique(av$region)

```


```{r echo = FALSE}

av_total_org <- subset(av, av$region == "TotalUS" & av$type == "organic")
av_total_con <- subset(av, av$region == "TotalUS" & av$type == "conventional")
av_total <- subset(av, av$region == "TotalUS")

region_list <- list('GreatLakes','Midsouth', 'Northeast','NorthernNewEngland','West', 'SouthCentral', 'Southeast','Plains', 'West')
av_regional_org <- subset(av, av$region %in% region_list & av$type == "organic")
av_regional_con <- subset(av, av$region %in% region_list & av$type == "conventional")
av_regional <- subset(av, av$region %in% region_list)

region_list <- list('GreatLakes','Midsouth', 'Northeast','NorthernNewEngland','West', 'SouthCentral', 'Southeast','Plains', 'West', 'TotalUS')
av_city_org <- subset(av, !(av$region %in% region_list) & av$type == "organic")
av_city_con <- subset(av, !(av$region %in% region_list) & av$type == "conventional")
av_city <- subset(av, !(av$region %in% region_list))


```



## Explore


### What happened to avocado prices and consumption in 2017?

```{r echo = FALSE, warning = FALSE}

ap1_org <- ggplot(aes(y = AveragePrice), data = subset(av, type == 'organic')) +
  geom_boxplot(fill = 'orange') + 
  scale_y_continuous(breaks = seq(0,3.5,.25)) +
  ylab('Average Price') +
  ggtitle('Organic Avocado Boxplot')

ap2_org <- ggplot(aes(x = AveragePrice), data = subset(av, type == 'organic')) +
  geom_histogram(fill = 'orange') +
  scale_y_continuous(breaks = seq(0,1000,100)) + 
  scale_x_continuous(breaks = seq(0,3.25,.5)) +
  ggtitle('Organic Avocado Histogram')

ap1_con <- ggplot(aes(y = AveragePrice), data = subset(av, type == 'conventional')) +
  geom_boxplot(fill = 'orange') + 
  scale_y_continuous(breaks = seq(0,2.5,.25)) +
  ylab('Average Price') +
  ggtitle('Conventional Avocado Boxplot')

ap2_con <- ggplot(aes(x = AveragePrice), data = subset(av, type == 'conventional')) +
  geom_histogram(fill = 'orange') +
  scale_y_continuous(breaks = seq(0,800,100)) +
  scale_x_continuous(breaks = seq(0,2.25,.25)) +
  ggtitle('Conventional Avocado Histogram')

suppressMessages(grid.arrange(ap1_org, ap2_org, ap1_con, ap2_con, ncol = 2)) 

```

We can see in the chart above that the median price of organic avocados is ~50
cents higher than conventional avocados. The price distributions for both
organic and conventional avocados are normally distributed.

People choose to buy organic avocados over conventional for a variety of
reasons, but some of the reasons contribute to an increase in prices. There is
typically a higher production cost associated with organic foods and the
production and distribution chains are not nearly as robust. For these reasons
and more, we should expect that the organic avocados are in general more
expensive.

Now let's take a look at the average avocado price over time and see if there
are any trends.

```{r echo = FALSE}

ggplot(aes(x = Date, y = AveragePrice, color = type), data = subset(av, region != 'TotalUS')) +
  geom_line(stat = 'summary', fun.y = mean, size = 1) +
  ggtitle('Average Price by Avocado Type Over Time')   

p2 <- ggplot(aes(x = Date, y = Total.Volume), data = subset(av, type == 'conventional' & region != 'TotalUS')) +
  geom_line(stat = 'summary', fun.y = mean) +
  stat_smooth(method = 'lm', color = 'blue') +
  scale_y_continuous(breaks = seq(0,2000000, 250000)) +
  scale_x_datetime(date_breaks = '6 months') +
  ggtitle('Average Regional Conventional Avocado Sales Over Time')

p3 <- ggplot(aes(x = Date, y = Total.Volume), data = subset(av, type == 'organic' & region != 'TotalUS')) +
  geom_line(stat = 'summary', fun.y = mean) +
  stat_smooth(method = 'lm', color = 'blue') +
  scale_x_datetime(date_breaks = '6 months') +
  ggtitle('Average Regional Organic Avocado Sales over Time')

grid.arrange(p2,p3,ncol = 1)

```

We can see in these graphs that there is a cyclical trend that has emerged for
the price of avocados regardless of type. The average price of avocados dip
around the new year and are most expensive during the summer. This also
corresponds to a yearly increase in avocado consumption around the new year.

Additionally, we can see that the total volume of conventional avocado sales has
remained relatively level but there is a steady increase in average organic
avocado sales. Even with the increase in organic avocado sales, the conventional
avocado sales volume still dwarfs that of the organic variety.

Outside of the continuing increase in organic avocado sales, it does not appear
that the price or consumption of avocados in general changed notably during
2017.

### How do the prices of avocados vary by region?

```{r echo = FALSE}

ggplot(aes(x = reorder(region, AveragePrice), y = AveragePrice, fill = type), data = av) +
  geom_bar(stat = 'summary', fun.y = mean, position = 'dodge') +
  xlab('City') +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  ggtitle('Average Avocado Price by Region') +
  scale_y_continuous(breaks=seq(0,2.5,.25)) +
  coord_flip()

```

Reported statistics say that the majority of avocados consumed in the U.S. are
produced either in Mexico or California. Based on that information, I would
expect that regions along the southern and western US borders have the lowest
distribution costs and therefore lower prices.

```{r echo = FALSE}

#https://stackoverflow.com/a/25664367
p1 <- ggplot(aes(x = reorder(region, -AveragePrice), y = AveragePrice), data = av_regional_con) +
  geom_bar(stat = 'summary', fun.y = mean) +
  xlab('Region') +
  ggtitle('Average Price of Avocados by Region')

p2 <- ggplot(aes(x = reorder(region, -Total.Volume), y = Total.Volume), data = av_regional_con) +
  geom_bar(stat = 'summary', fun.y = mean) +
  xlab('Region') +
  ylab('Average Volume') +
  scale_y_continuous(breaks = seq(0,6000000,1000000)) +
  ggtitle('Average Sales Volume of Avocado by Region')

grid.arrange(p1,p2, ncol = 1)


```

Here we can see the distribution costs reflected in the average price by region.
The South-Central, West, and and South-East regions have the lowest average
price whereas the North-East and New England regions have the highest. I would
also guess that avocado availability is less sensitive to the seasonal changes
in the south and west. This combined with the price discrepancy probably
contributes significantly to these regions high average sales volumes.


## Random Analysis


Sales of Conventional Avocados by Size Classification

```{r echo = FALSE}

p1 <- ggplot(aes(x = Date, y = X4046), data = subset(av, region == 'TotalUS' & type == 'conventional')) +
  geom_line(stat = 'summary', fun.y = mean, size = 1) +
  ggtitle('Average Sales of X4046 (small) Conventional Avocados Type Over Time')

p2 <- ggplot(aes(x = Date, y = X4225), data = subset(av, region = 'TotalUS' & type == 'conventional')) +
  geom_line(stat = 'summary', fun.y = mean, size = 1) +
  ggtitle('Average Sales of X4225 (medium) Conventional Avocados Type Over Time')   

p3 <- ggplot(aes(x = Date, y = X4770), data = subset(av, region = 'TotalUS' & type == 'conventional')) +
  geom_line(stat = 'summary', fun.y = mean, size = 1) +
  ggtitle('Average Sales of X4770 (large) Conventional Avocados Type Over Time')

grid.arrange(p1,p2,p3, ncol = 1)

```

```{r}

p1 <- ggplot(aes(x = Date, y = X4046), data = subset(av, region == 'TotalUS' & type == 'organic')) +
  geom_line(stat = 'summary', fun.y = mean, size = 1) +
  ggtitle('Average Sales of X4046 (small) Organic Avocados Type Over Time')

p2 <- ggplot(aes(x = Date, y = X4225), data = subset(av, region = 'TotalUS' & type == 'organic')) +
  geom_line(stat = 'summary', fun.y = mean, size = 1) +
  ggtitle('Average Sales of X4225 (medium) Organic Avocados Type Over Time')   

p3 <- ggplot(aes(x = Date, y = X4770), data = subset(av, region = 'TotalUS' & type == 'organic')) +
  geom_line(stat = 'summary', fun.y = mean, size = 1) +
  ggtitle('Average Sales of X4770 (large) Organic Avocados Type Over Time')

grid.arrange(p1,p2,p3, ncol = 1)


```


